### 2013-10-04 chi2014/qualitative.tex
more revisions in progress

In addition to modelling data and confirming our hypotheses, we performed a qualitative analysis.
We selected 30 brainstorming runs (5 from each condition). The primary author applied open coding to identify characteristics of responses and runs that are independent of the uniqueness and originality measures observed above. In particular, we identified groups of ideas within runs that shared characteristics. These characteristics, in addition to temporal order, form the basis for what we call \emph{strategies}: repeatedly observed classes of ideas that are generated by some participants but not others.
Furthermore, we use the presence or lack of presence of strategies to define \emph{personas}.

@@

A \emph{strategy} is a class of ideas repeatedly observed in groups, though not neccessarily consecutively, within brainstorming runs. These ideas are defined by some common characteristics and may or may not be further defined by temporal order. Not all participants utilized all strategies in their brainstorming runs.

We identified four categories of strategy. The first, \emph{feature exploration}, is specific to the MP3 player problem presented in the paper. This category entails focussing on specific features a device possesses and leveraging them as the basis for generating new ideas. The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying six(?) different types of riffing exployed. \emph{Time-saving} strategies are those that allow the participant to generate at a faster rate. Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.

\subsubsection{Feature exploration}

One category of strategies employed was \emph{feature exploration}. An MP3 player has many features that distinguish it from any other electronic device of the same physical proportions: audio output (and potentially input), data storage, a screen, general purpose computing abilities (apps), connectivity (WiFi, USB), a camera, a unique appearance, GPS.
Each of these features forms the basis for a strategy of idea generation.
Participants would choose a feature and generate ideas that leveraged that feature in various ways. For example, here are several answers from a participant who explored audio input and output:

@@

While it might be expected that all participants would employ this strategy to some degree, many treated the MP3 player as a generic electronic device of equivalent size. Runs were commonly composed of answers that could apply to almost any un-used household item:

@@

Runs that used a feature exploration strategy produce responses that seem intuitively better than those that do not. It is satisfying to have a use for a device that "fits" the device as closely as possible, so that none of the functionality is "wasted".

@@

\emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:

@@

Another, more common type of riffing is \emph{exhaustive mutation}. Under this strategy, a participant repeats a response nearly verbatim many times, each time replacing a single word or concept in such a way that no new ideas are formed. For example:

@@

\emph{Exploratory riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed. This can be chained many times in a kind of idea "telephone game" such that the final ideas no longer resemble the original source. For example:

\begin{itemize}
    \item Jukebox music selector in bars
    \item Commercials in bar bathrooms (retains bar element)
    \item Tapper handles for beer (retain alcohol element)
\end{itemize}

Finally, \emph{Follow-up riffing} occurs when a participant's response assumes that the solution in one of their previous responses has already been implemented. For example:

>>>>

In addition to modelling data and confirming our hypotheses, we performed a qualitative analysis. The primary goal was to identify trends  that were repeated between participants. These trends can then be used to inform and justify interventions to change the types of ideas produced.

We selected 30 brainstorming runs (5 randomly from each condition). The primary author applied an open coding strategy, flagging ideas within runs that shared common characteristics, such as length, wording, or core problem solution.

In particular, we identify classes of ideas within runs, often temporally close, that hold common characteristics. For each class that was observed across multiple participants, we infer the existence of a \emph{strategy}: a plan or rule that aids in generating ideas. We describe the observed strategies in detail below.
Furthermore, we use the presence or lack of presence of strategies to define \emph{personas}.

@@

A \emph{strategy} is a plan, rule, or idea that is used by a participant to generate brainstorming responses.
Without debriefing participants, we cannot infer the goal of a strategy, but define it in terms of its observable effects.
These effects are repeated, cross-participant presence of classes of similar ideas.
Not all participants utilized all strategies in their brainstorming runs.

We identified X \emph{orthogonal} strategy categories - each category of strategies can be employed by a participant with or without any of the other categories.
The first, \emph{problem scoping}, is a process in which a participant chooses a limited implication or component of the problem and provides multiple specific solutions.
The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying X different types of riffing exployed.
Finally, \emph{time-saving} strategies increase the rate of idea generation.
%Finally, the \emph{humour} strategy emphasizes generating ideas that are amusing rather than practical.

\subsubsection{Problem scoping}

To complete the brainstorming task, participants had to generate some number of uses for a partially broken MP3 player. Under the problem scoping strategy, this problem can reduced in scope to one of many sub-problems: brainstorm uses for \emph{an audio output device}; brainstorm uses for \emph{an electronic screen}; brainstorm uses for \emph{a hard drive}; and so on. The participant chooses one sub-problem at a time and generates solutions, each of which is also a solution to the original problem. For example, the following are generated by a participant focussing on the audio output capabilities of an MP3 player:

@@

The previous examples of problem scoping are examples of emphasizing specific details of the problem. Problem scoping can also entail neglecting details. For example, the following ideas from a participant apply to almost any physical object, not just the MP3 player described:

@@

We refer to the former, detail-oriented kind of problem scoping as \emph{scoping inward} and the alternative as \emph{scoping outward}. Problem scoping is employed by all participants, but the type of scoping (inwards vs outward) is a useful discriminant.

@@

\emph{Generalization riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:

@@

Another type of riffing is \emph{exhaustive mutation}. Under this strategy, a participant repeats a response nearly verbatim many times, each time replacing a single word or concept in such a way that no new ideas are formed. For example:

@@

\emph{Hold riffing} is a strategy in which participants hold at least one element of a previous response constant when generating a new response, such that a new idea is formed. This can be thought of as analogous to the game of \emph{draw poker}, in which players hold some cards while receiving new ones, with the intent of building a winning hand complementing the held cards.
For example:

\begin{itemize}
    \item Jukebox music selector in bars
    \item Commercials in bar bathrooms (hold the bar element)
    \item Tapper handles for beer (hold the alcohol element)
\end{itemize}

Finally, \emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented. The new ideas relies on the implications or consequences of the previous. For example:
